Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2019-01-28 writing/ehp_methods.tex
Work some on formatting

All counties in the eastern half of the \ac{US}
(Figure~\ref{fig:ivanexposure}), as of the~2010 \ac{US} Decennial Census, were
included in the study. The study covered all tracked storms in the \ac{HURDAT2}
\citep{landsea2013} that came within~250 \si{\kilo\metre} of at least one study
county between~1988 and~2015 (Figure~S2).  \ac{HURDAT2} is a post-storm
assessment conducted by the \ac{US} \ac{NHC} and incorporates data from a
variety of sources, including satellite data and, when available, aircraft
reconnaissance data~\citep{landsea2013, jarvinen1988}. These data typically
give measurements of the tropical cyclone center's location at~6-\si{\hour} intervals
at synoptic times (i.e., 6:00~am, 12:00~pm, 6:00~pm, and 12:00~am \ac{UTC});
some landfalling tropical cyclones have an additional observation at the time
of landfall~\citep{landsea2013}.

@@

For each storm, the distance of closest approach was measured for every study
county.  Storm center track location data from \ac{HURDAT2} were interpolated
using natural cubic splines to~15-\si{\minute} intervals~\citep{hurricaneexposure}.
The distance between each county's population mean center, as assessed by the
2010 Decennial Census~\citep{countycenters}, and each~15-\si{\minute} interpolated
storm center location was measured using Great Circle (WGS84 ellipsoid)
distance~\citep{bivand2013applied}. The distance-based exposure metric was
based on the minimum value of this storm center-to-county distance throughout
the period the storm was tracked. The time of closest approach for each
county-storm pairing was also recorded for use assessing rain-, flood-, and
tornado-based metrics. To allow matching with data based on local time, times
were converted from \ac{UTC} to local time~\citep{countytimezones}.

@@

For the rain-based exposure metric, cumulative rainfall was estimated for the
period from two days before to one day after the date of a tropical cyclone's
closest approach to a county. Daily county-level rainfall was determined by
aggregating hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
precipitation data files~\citep{rui2013nldas}. The \ac{NLDAS-2} data integrate
satellite-based and land-based monitoring, applying a land-surface model to
create a reanalysis dataset that is spatially and temporally complete across
the continental \ac{US}~\citep{rui2013nldas, alhamdan2014environmental}. To
aggregate to county-level, hourly data at each grid point were summed to create
a daily rainfall total, and these grid point rainfall totals were then averaged
across all grid points within a county's~1990 \ac{US} Census boundaries to
create a county-level daily total~\citep{alhamdan2014environmental, cdcwonder}.
This county-level precipitation data are publicly available through the \ac{US}
\ac{CDC}'s \ac{WONDER} database~\citep{cdcwonder, alhamdan2014environmental}.
The rainfall-based exposure metric was only available for tropical cyclones
through~2011, the period for which the county-aggregated reanalysis rainfall
data were available~\citep{cdcwonder, alhamdan2014environmental}.

To assess the rainfall estimates from the reanalysis data used in analysis in
the the main text, we compared cumulative rainfall estimates based on this
reanalysis rainfall data to rainfall observations from ground-based weather
stations in nine sample study counties, selected as counties with data
available from multiple stations. In these nine counties, observed rainfall was
collected for all storms,~1988\,--\,2011, for the same period of two days
before to one day after each storm's closest approach, using all stations with
available daily data in the Global Historical Climatology Network throughout
1988\,--\,2011~\citep{menne2012overview, rnoaa, countyweather}. Agreement
between measured storm-related rainfall between these two data sources was
assessed by Spearman's rank correlation.

@@

Ground-based observations of wind speed are problematic during tropical
cyclones, as instruments often fail at high wind speed, while reanalysis data,
often available at hourly or higher resolution, can be too smooth to capture
wind extremes associated with a tropical cyclone. Therefore, to create a
dataset of county-level sustained winds during historical tropical cyclones, we
modeled maximum sustained wind speeds at each county's population mean
center~\citep{countycenters}. \ac{HURDAT2} Best Track data were interpolated
from~6-\si{\hour} reported values to~15-\si{\minute} increments using natural cubic
splines~\citep{stormwindmodel}, and then sustained wind speed was modeled for
each county center at each~15-\si{\minute} increment using a double exponential wind
speed model~\citep{willoughby2006parametric} to estimate maximum ground-level
sustained wind speed at each county center~\citep{stormwindmodel}. Asymmetry in
wind speeds around the tropical cyclone center was incorporated into the wind
speed model~\citep{phadke2003modeling, stormwindmodel}. The maximum value of
this modeled wind speed was determined for each county as the maximum windspeed
across the~15-\si{\minute} incremented modeled values throughout the tropical
cyclone.

To validate the metric of county-level wind speed used in the main text, we
compared this wind metric to estimated wind radii from \ac{HURDAT2}, available
for storms in our study from~2004~\citep{landsea2013}. \ac{HURDAT2} gives
estimates of the maximum radii to winds of~64,~50, and~34 \si{\knot} in each four
quadrants of a storm (i.e., distance from the storm's center to the furthest
point in a quadrant with winds at or above that speed) at each~6-\si{\hour} storm
observation.

@@

To identify flood- and tornado-based exposures to tropical cyclones in \ac{US}
counties, we used event listings from the National Oceanic and Atmospheric
Administration (NOAA)'s Storm Event Database~\citep{stormevents}. For each
tropical cyclone, we identified all events with event types related to flooding
(``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's track
and with a start date within a five-day window centered on the date of the
tropical cyclone's closest approach to the
county~\citep{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
events in this database were reported by county \ac{FIPS} code. ``Coastal
Flood" events were reported by forecast zone; for these, the event was matched
to the appropriate county if possible~\citep{noaastormevents}. While this
database has recorded storm data since~1950, its coverage changed substantially
in~1996 to cover a wider variety of storm events~\citep{stormevents}. We
therefore only considered these metrics of hurricane exposure for tropical
cyclones in~1996 and later.

As validation for the flood data, we compared streamflow data from \ac{US}
Geological Survey gages~\citep{usgsgages, countyfloods, dataRetrieval} in nine
counties with tropical cyclone flood classifications based on the NOAA Storm
Events Database (Figure~\ref{fig:floodcomparison}).

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary,
since these metrics were based on whether an event was listed in the NOAA Storm
Events database or not.  For the other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
assessed county-level tropical cyclone exposure in the eastern \ac{US} based on
each exposure metric for all land-falling or near-land Atlantic basin tropical
cyclones. Depending on available exposure data, this assessment included some
or all of the period from~1988 to~2015, a period with~136 tropical cyclones
that made \ac{US} landfall or passed within~250 \si{\kilo\metre} (155 mi) of at
least one \ac{US} county (Figure~S2).

We measured agreement between exposure metrics in the classification of a
county as exposed or unexposed to a specific tropical cyclone. For this
measurement, we calculated the within-storm Jaccard
index~\citep{jaccard1901distribution, jaccard1908nouvelles} between every
pair-wise combination of hazard metrics.  The Jaccard index~($J_s$) measures
similarity between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical
cyclone~$s$ as the proportion of counties in which both of the metrics classify
the county as exposed out of all counties classified as exposed by at least one
of the metrics:

@@

The outcomes of disasters depend both on the geophysical forces of the disaster
as well as on the vulnerability of the society living in the geographical areas
affected by those forces.~\citep{chakraborty2005population,
anderson2003community, cutter1996vulnerability} As a case study of how
differences in tropical cyclone exposure assessments across different hazard
metrics influence estimates of physical exposures of susceptible populations to
tropical cyclones, we investigated how the use of different metrics influenced
estimates of which eastern \ac{US} counties have the highest expected average
exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
This subpopulation was selected for this analysis since it may be particularly
susceptible to health impacts from tropical cyclone exposure, especially
through the pathway of storm-associated power outages and evacuations. We
collected data on the number of electricity-dependent Medicare beneficiaries in
each study county from the U.S. Department of Health \& Human Service's emPower
Map~2.0~\citep{empower}. We then calculated the physical exposure of the
electricity-dependent Medicare population in each county, based on tropical
cyclone assessments using each hazard metric,
following~\citep{peduzzi2009assessing}:

@@

\noindent where~$E_c$ is the average yearly physical exposure among
electricity-dependent Medicare beneficiaries in county~$c$ to tropical cyclone
exposures based on a given metric,~$F_c$ is the estimated yearly expected
frequency of tropical cyclone exposures in county~$c$ based on that metric,
and~$P_c$ is the size of the electricity-dependent Medicare population in
county~$c$, as of July~2017. When combined with estimates of vulnerability of a
population to a natural hazard, such measurements of physical exposure can be
used to calculate risk of human losses from the
hazard~\citep{peduzzi2009assessing}.

>>>>

All counties in the eastern half of the \ac{US}
(Figure~\ref{fig:ivanexposure}), as of the~2010 \ac{US} Decennial Census, were
included in the study. The study covered all tracked storms in the \ac{HURDAT2}
\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
study county between~1988 and~2015 (Figure~S2\todo{Check fig ref}).
\ac{HURDAT2} is a post-storm assessment conducted by the \ac{US} \ac{NHC} and
incorporates data from a variety of sources, including satellite data and, when
available, aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.
These data typically give measurements of the tropical cyclone center's
location at~6-\si{\hour} intervals at synoptic times (i.e., 6:00~am, 12:00~pm,
6:00~pm, and 12:00~am \ac{UTC}); some landfalling tropical cyclones have an
additional observation at the time of landfall~\parencite{landsea2013}.

@@

For each storm, the distance of closest approach was measured for every study
county.  Storm center track location data from \ac{HURDAT2} were interpolated
using natural cubic splines to~15-\si{\minute} intervals~\parencite{hurricaneexposure}.
The distance between each county's population mean center, as assessed by the
2010 Decennial Census~\parencite{countycenters}, and each~15-\si{\minute} interpolated
storm center location was measured using Great Circle (WGS84 ellipsoid)
distance~\parencite{bivand2013applied}. The distance-based exposure metric was
based on the minimum value of this storm center-to-county distance throughout
the period the storm was tracked. The time of closest approach for each
county-storm pairing was also recorded for use assessing rain-, flood-, and
tornado-based metrics. To allow matching with data based on local time, times
were converted from \ac{UTC} to local time~\parencite{countytimezones}.

@@

For the rain-based exposure metric, cumulative rainfall was estimated for the
period from two days before to one day after the date of a tropical cyclone's
closest approach to a county. Daily county-level rainfall was determined by
aggregating hourly~1/8\si{\degree} gridded reanalysis data from the \ac{NLDAS-2}
precipitation data files~\parencite{rui2013nldas}. The \ac{NLDAS-2} data integrate
satellite-based and land-based monitoring, applying a land-surface model to
create a reanalysis dataset that is spatially and temporally complete across
the continental \ac{US}~\parencite{rui2013nldas, alhamdan2014environmental}. To
aggregate to county-level, hourly data at each grid point were summed to create
a daily rainfall total, and these grid point rainfall totals were then averaged
across all grid points within a county's~1990 \ac{US} Census boundaries to
create a county-level daily total~\parencite{alhamdan2014environmental, cdcwonder}.
This county-level precipitation data are publicly available through the \ac{US}
\ac{CDC}'s \ac{WONDER} database~\parencite{cdcwonder, alhamdan2014environmental}.
The rainfall-based exposure metric was only available for tropical cyclones
through~2011, the period for which the county-aggregated reanalysis rainfall
data were available~\parencite{cdcwonder, alhamdan2014environmental}.

To assess the rainfall estimates from the reanalysis data used in analysis in
the the main text, we compared cumulative rainfall estimates based on this
reanalysis rainfall data to rainfall observations from ground-based weather
stations in nine sample study counties, selected as counties with data
available from multiple stations. In these nine counties, observed rainfall was
collected for all storms,~1988\,--\,2011, for the same period of two days
before to one day after each storm's closest approach, using all stations with
available daily data in the Global Historical Climatology Network throughout
1988\,--\,2011~\parencite{menne2012overview, rnoaa, countyweather}. Agreement
between measured storm-related rainfall between these two data sources was
assessed by Spearman's rank correlation.

@@

Ground-based observations of wind speed are problematic during tropical
cyclones, as instruments often fail at high wind speed, while reanalysis data,
often available at hourly or higher resolution, can be too smooth to capture
wind extremes associated with a tropical cyclone. Therefore, to create a
dataset of county-level sustained winds during historical tropical cyclones, we
modeled maximum sustained wind speeds at each county's population mean
center~\parencite{countycenters}. \ac{HURDAT2} Best Track data were interpolated
from~6-\si{\hour} reported values to~15-\si{\minute} increments using natural cubic
splines~\parencite{stormwindmodel}, and then sustained wind speed was modeled for
each county center at each~15-\si{\minute} increment using a double exponential wind
speed model~\parencite{willoughby2006parametric} to estimate maximum ground-level
sustained wind speed at each county center~\parencite{stormwindmodel}. Asymmetry in
wind speeds around the tropical cyclone center was incorporated into the wind
speed model~\parencite{phadke2003modeling, stormwindmodel}. The maximum value of
this modeled wind speed was determined for each county as the maximum windspeed
across the~15-\si{\minute} incremented modeled values throughout the tropical
cyclone.

To validate the metric of county-level wind speed used in the main text, we
compared this wind metric to estimated wind radii from \ac{HURDAT2}, available
for storms in our study from~2004~\parencite{landsea2013}. \ac{HURDAT2} gives
estimates of the maximum radii to winds of~64,~50, and~34 \si{\knot} in each four
quadrants of a storm (i.e., distance from the storm's center to the furthest
point in a quadrant with winds at or above that speed) at each~6-\si{\hour} storm
observation.

@@

To identify flood- and tornado-based exposures to tropical cyclones in \ac{US}
counties, we used event listings from the National Oceanic and Atmospheric
Administration (NOAA)'s Storm Event Database~\parencite{stormevents}. For each
tropical cyclone, we identified all events with event types related to flooding
(``Flood", ``Flash Flood", ``Coastal Flood") and tornadoes (``Tornado") and
that occurred in a county within~500 \si{\kilo\metre} of the tropical cyclone's track
and with a start date within a five-day window centered on the date of the
tropical cyclone's closest approach to the
county~\parencite{hurricaneexposuredata}. ``Flood", ``Flash Flood" and ``Tornado"
events in this database were reported by county \ac{FIPS} code. ``Coastal
Flood" events were reported by forecast zone; for these, the event was matched
to the appropriate county if possible~\parencite{noaastormevents}. While this
database has recorded storm data since~1950, its coverage changed substantially
in~1996 to cover a wider variety of storm events~\parencite{stormevents}. We
therefore only considered these metrics of hurricane exposure for tropical
cyclones in~1996 and later.

As validation for the flood data, we compared streamflow data from \ac{US}
Geological Survey gages~\parencite{usgsgages, countyfloods, dataRetrieval} in nine
counties with tropical cyclone flood classifications based on the NOAA Storm
Events Database (Figure~\ref{fig:floodcomparison}).

@@

Two of the exposure metrics (flood- and tornado-based) were inherently binary,
since these metrics were based on whether an event was listed in the NOAA Storm
Events database or not.  For the other exposure metrics, each county was
classified as exposed to a tropical cyclone based on whether the exposure
metric exceeded a certain threshold (Table~\ref{tab:exposuremetrics}). We
assessed county-level tropical cyclone exposure in the eastern \ac{US} based on
each exposure metric for all land-falling or near-land Atlantic basin tropical
cyclones. Depending on available exposure data, this assessment included some
or all of the period from~1988 to~2015, a period with~136 tropical cyclones
that made \ac{US} landfall or passed within~250 \si{\kilo\metre} (155 mi) of at
least one \ac{US} county (Figure~S2\todo{Check fig ref}).

We measured agreement between exposure metrics in the classification of a
county as exposed or unexposed to a specific tropical cyclone. For this
measurement, we calculated the within-storm Jaccard
index~\parencite{jaccard1901distribution, jaccard1908nouvelles} between every
pair-wise combination of hazard metrics.  The Jaccard index~($J_s$) measures
similarity between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical
cyclone~$s$ as the proportion of counties in which both of the metrics classify
the county as exposed out of all counties classified as exposed by at least one
of the metrics:

@@

The outcomes of disasters depend both on the geophysical forces of the disaster
as well as on the vulnerability of the society living in the geographical areas
affected by those forces.~\parencite{chakraborty2005population,
anderson2003community, cutter1996vulnerability} As a case study of how
differences in tropical cyclone exposure assessments across different hazard
metrics influence estimates of physical exposures of susceptible populations to
tropical cyclones, we investigated how the use of different metrics influenced
estimates of which eastern \ac{US} counties have the highest expected average
exposures of electricity-dependent Medicare beneficiaries to tropical cyclones.
This subpopulation was selected for this analysis since it may be particularly
susceptible to health impacts from tropical cyclone exposure, especially
through the pathway of storm-associated power outages and evacuations. We
collected data on the number of electricity-dependent Medicare beneficiaries in
each study county from the U.S. Department of Health \& Human Service's emPower
Map~2.0~\parencite{empower}. We then calculated the physical exposure of the
electricity-dependent Medicare population in each county, based on tropical
cyclone assessments using each hazard metric,
following~\parencite{peduzzi2009assessing}:

@@

\noindent where~$E_c$ is the average yearly physical exposure among
electricity-dependent Medicare beneficiaries in county~$c$ to tropical cyclone
exposures based on a given metric,~$F_c$ is the estimated yearly expected
frequency of tropical cyclone exposures in county~$c$ based on that metric,
and~$P_c$ is the size of the electricity-dependent Medicare population in
county~$c$, as of July~2017. When combined with estimates of vulnerability of a
population to a natural hazard, such measurements of physical exposure can be
used to calculate risk of human losses from the
hazard~\parencite{peduzzi2009assessing}.

### 2019-01-28 writing/ehp_results.tex
Work some on formatting

Figure~\ref{fig:ivanexposure} gives an example of how sensitive county-level
tropical cyclone exposure assessment is to the choice of metric, using as an
example Hurricane Ivan in 2004, which was record-breaking for its
duration as a major hurricane~\citep{franklin2006atlantic}. When county-level
exposure was determined based on distance, the counties assessed as exposed
followed the tropical cyclone's track, including counties well inland. For the
wind metric, only counties near two of the tropical cyclone's landfalls were
assessed as exposed. For rain- and flood-based metrics, however, exposure
extended to the left of the track, including counties as far north as New York
and Connecticut, while for the tornado metric, exposed counties tended to be to
the right of the track and included several counties in central North Carolina,
South Carolina, and Georgia that were not identified as exposed to Ivan based
on any other metric.  Figure~S3 provides similar maps for four other example
tropical cyclones (selected because they exposed many \ac{US}  counties based on
at least one metric), and Figure~S4 provides four examples of tropical cyclones
in which the disagreement in exposure assessment between rain- and wind-based
metrics was particularly notable.

@@

There were, however, a few exceptions\,---\,tropical cyclones in which similar
counties were determined to be exposed to the tropical cyclone for two or more
of the metrics considered.  For Floyd in~1999 (Figure~S3) and Irene in~2011,
for example, county-level classification agreed moderately to well (Jaccard
index of approximately~0.5\,--\,0.8) for all pairs of exposure metrics except
those including the tornado-based metric. These tropical cyclones both made
their first \ac{US}  landfall in North Carolina at minor hurricane windspeeds
(Category~2 and~1, respectively) and then skimmed the eastern coast of the
\ac{US} north through New England, bringing large rainfall to much of the
eastern coast from North Carolina north and causing extensive inland flooding
in North Carolina (Floyd) and New England (Irene)~\citep{avila2013atlantic,
lawrence2000atlantic}. For another set of tropical cyclones (e.g., Lee in~2011
[Figure~S3], Ernesto in~2006, and Bertha in~1996), there was moderate to good
agreement for all pairwise combinations of distance, rain, and wind, but poor
agreement for all other combinations of metrics. Most of these tropical
cyclones either followed the eastern coastline north of North Carolina (Bertha
in~1996, Charley in~2004, Gaston in~2004, Barry in~2007, and Hanna in~2008), in
a pattern similar to Hurricanes Floyd in~1999 and Irene in~2011, or made
landfall in or near the Florida panhandle and cut across to exit into the
Atlantic around North Carolina (Josephine in~1996, Earl in~1998, Alberto
in~2006). However, these two sets of tropical cyclones represented the minority
of all tropical cyclones considered; for most tropical cyclones, there was low
overlap in the counties determined to be exposed to the tropical cyclones
between most pairings of metrics.

@@

We conducted a case study to investigate how these differences in exposure
assessment across the five metrics considered could strongly influence
health-related studies based on physical exposure to tropical cyclones. For
this example, we compared estimates of which \ac{US} counties have the largest
expected physical exposure to tropical cyclones among a susceptible
subpopulation, Medicare beneficiaries who are reliant on electricity for
medical equipment (Figure~S6)~\citep{empower}.  We found
that a few counties (Miami-Dade County, FL; Harris County, TX) were ranked in
the top three counties of physical exposure among electricity-dependent
Medicare beneficiaries for almost all exposure metrics
(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
when using different exposure metrics. For example, Florida counties dominated
the lists of top exposed counties based on wind- and tornado-based metrics,
while over half of the counties for the flood-based metric were in Mid-Atlantic
states (PA, NY).

@@

To assist with future tropical cyclone impact studies, we created and published
open source software that includes the hazard-specific, county-level tropical
cyclone data in this paper, as well as tools to explore and map that 
data~\citep{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
data for the distance-, rain-, and wind-based metrics, allowing users to also explore
different thresholds of these metrics when assessing exposure.

We investigated how well these data correspond with data from other available
sources (Figures~S7\,--\,11). Within the nine sample counties, storm-related
rainfall from the reanalysis precipitation dataset was well-correlated with
station-based rainfall measurements, although the reanalysis data did tend to
somewhat underestimate rainfall totals in storms with extremely high rainfall
(Figure~\ref{fig:raincomparison}). A distance constraint (county population
mean center within~500 \si{\kilo\metre} of the storm's closest approach) was
necessary for assessing storm exposure based on this metric to ensure that
rains unrelated and far from the storm track were not misattributed to storm
exposures. We found that this distance constraint was typically adequately
large to capture storm-related rain, although in rare examples of exceptionally
large storms (e.g., Hurricane Ike in 2008) or storms for which tracking was
stopped at extratropical transition (e.g., Tropical Storm Lee in 2011), some
storm-related rains were missed because of this distance constraint
(Figure~\ref{fig:rainexamples}).  The modeled wind speeds generally agreed well
with the \ac{HURDAT2} wind radii (Figure~\ref{fig:windcomparison}), with
disagreement mostly limited to large storms that were characterized by high
winds well inland (e.g., Hurricanes Ike in 2008 and Sandy in 2012;
Figure~\ref{fig:windexamples}).  Stream gage data generally indicated higher
discharge in these counties on dates identified as flood events, particularly
in counties with more reporting stream gages (e.g., Harris County, TX, with~10
reporting stream gages; Figure~\ref{fig:floodcomparison}), although these two
flooding data sources were not completely consistent.

While generally in agreement with data from other
sources, there are a few limitations. The wind data are based on modeled,
rather than observed, values, and while the modeled wind data generally agree
well with post-analysis maximum wind radii from the revised Atlantic hurricane
database \ac{HURDAT2}~\citep{landsea2013} (Figure~S7), for a few tropical cyclones
it did not (e.g., Figure~S10). The rainfall data are from re-analysis data,
which are generally well-correlated with observed ground-based station data but
sometimes oversmooth extreme measurements (Figure~S8).

>>>>

Figure~\ref{fig:ivanexposure} gives an example of how sensitive county-level
tropical cyclone exposure assessment is to the choice of metric, using as an
example Hurricane Ivan in 2004, which was record-breaking for its duration as a
major hurricane~\parencite{franklin2006atlantic}. When county-level exposure
was determined based on distance, the counties assessed as exposed followed the
tropical cyclone's track, including counties well inland. For the wind metric,
only counties near two of the tropical cyclone's landfalls were assessed as
exposed. For rain- and flood-based metrics, however, exposure extended to the
left of the track, including counties as far north as New York and Connecticut,
while for the tornado metric, exposed counties tended to be to the right of the
track and included several counties in central North Carolina, South Carolina,
and Georgia that were not identified as exposed to Ivan based on any other
metric.  Figure~S3\todo{Check fig ref} provides similar maps for four other
example tropical cyclones (selected because they exposed many \ac{US}  counties
based on at least one metric), and Figure~S4\todo{Check fig ref} provides four
examples of tropical cyclones in which the disagreement in exposure assessment
between rain- and wind-based metrics was particularly notable.

@@

There were, however, a few exceptions\,---\,tropical cyclones in which similar
counties were determined to be exposed to the tropical cyclone for two or more
of the metrics considered.  For Floyd in~1999 (Figure~S3\todo{Check fig ref})
and Irene in~2011, for example, county-level classification agreed moderately
to well (Jaccard index of approximately~0.5\,--\,0.8) for all pairs of exposure
metrics except those including the tornado-based metric. These tropical
cyclones both made their first \ac{US}  landfall in North Carolina at minor
hurricane windspeeds (Category~2 and~1, respectively) and then skimmed the
eastern coast of the \ac{US} north through New England, bringing large rainfall
to much of the eastern coast from North Carolina north and causing extensive
inland flooding in North Carolina (Floyd) and New England
(Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}. For another set of
tropical cyclones (e.g., Lee in~2011 [Figure~S3\todo{Check fig ref}], Ernesto
in~2006, and Bertha in~1996), there was moderate to good agreement for all
pairwise combinations of distance, rain, and wind, but poor agreement for all
other combinations of metrics. Most of these tropical cyclones either followed
the eastern coastline north of North Carolina (Bertha in~1996, Charley in~2004,
Gaston in~2004, Barry in~2007, and Hanna in~2008), in a pattern similar to
Hurricanes Floyd in~1999 and Irene in~2011, or made landfall in or near the
Florida panhandle and cut across to exit into the Atlantic around North
Carolina (Josephine in~1996, Earl in~1998, Alberto in~2006). However, these two
sets of tropical cyclones represented the minority of all tropical cyclones
considered; for most tropical cyclones, there was low overlap in the counties
determined to be exposed to the tropical cyclones between most pairings of
metrics.

@@

We conducted a case study to investigate how these differences in exposure
assessment across the five metrics considered could strongly influence
health-related studies based on physical exposure to tropical cyclones. For
this example, we compared estimates of which \ac{US} counties have the largest
expected physical exposure to tropical cyclones among a susceptible
subpopulation, Medicare beneficiaries who are reliant on electricity for
medical equipment (Figure~S6\todo{Check fig ref})~\parencite{empower}.  We
found that a few counties (Miami-Dade County, FL; Harris County, TX) were
ranked in the top three counties of physical exposure among
electricity-dependent Medicare beneficiaries for almost all exposure metrics
(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
when using different exposure metrics. For example, Florida counties dominated
the lists of top exposed counties based on wind- and tornado-based metrics,
while over half of the counties for the flood-based metric were in Mid-Atlantic
states (PA, NY).

@@

To assist with future tropical cyclone impact studies, we created and published
open source software that includes the hazard-specific, county-level tropical
cyclone data in this paper, as well as tools to explore and map that 
data~\parencite{hurricaneexposure, hurricaneexposuredata}. This software includes continuous
data for the distance-, rain-, and wind-based metrics, allowing users to also explore
different thresholds of these metrics when assessing exposure.

We investigated how well these data correspond with data from other available
sources (Figures~S7\,--\,11\todo{Check fig refs}). Within the nine sample
counties, storm-related rainfall from the reanalysis precipitation dataset was
well-correlated with station-based rainfall measurements, although the
reanalysis data did tend to somewhat underestimate rainfall totals in storms
with extremely high rainfall (Figure~\ref{fig:raincomparison}). A distance
constraint (county population mean center within~500 \si{\kilo\metre} of the
storm's closest approach) was necessary for assessing storm exposure based on
this metric to ensure that rains unrelated and far from the storm track were
not misattributed to storm exposures. We found that this distance constraint
was typically adequately large to capture storm-related rain, although in rare
examples of exceptionally large storms (e.g., Hurricane Ike in 2008) or storms
for which tracking was stopped at extratropical transition (e.g., Tropical
Storm Lee in 2011), some storm-related rains were missed because of this
distance constraint (Figure~\ref{fig:rainexamples}).  The modeled wind speeds
generally agreed well with the \ac{HURDAT2} wind radii
(Figure~\ref{fig:windcomparison}), with disagreement mostly limited to large
storms that were characterized by high winds well inland (e.g., Hurricanes Ike
in 2008 and Sandy in 2012; Figure~\ref{fig:windexamples}).  Stream gage data
generally indicated higher discharge in these counties on dates identified as
flood events, particularly in counties with more reporting stream gages (e.g.,
Harris County, TX, with~10 reporting stream gages;
Figure~\ref{fig:floodcomparison}), although these two flooding data sources
were not completely consistent.

While generally in agreement with data from other sources, there are a few
limitations. The wind data are based on modeled, rather than observed, values,
and while the modeled wind data generally agree well with post-analysis maximum
wind radii from the revised Atlantic hurricane database
\ac{HURDAT2}~\parencite{landsea2013} (Figure~S7\todo{Check fig ref}), for a few
tropical cyclones it did not (e.g., Figure~S10\todo{Check fig ref}). The
rainfall data are from re-analysis data, which are generally well-correlated
with observed ground-based station data but sometimes oversmooth extreme
measurements (Figure~S8\todo{Check fig ref}).

### 2019-01-28 writing/ehp_tables.tex
Work some on formatting

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988--2015 & County population mean center within 100 \si{\kilo\metre} of storm track \\ 
Rain & 1988--2011 & County cumulative rainfall of 75 millimeters or more over the period from two days before to one day after the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\ 
Wind & 1988--2015 & Modeled maximum sustained wind speed at the county's population mean center 15 meters per second or higher during the storm\\ 
Flood & 1996--2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track \\
Tornado & 1996--2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500 \si{\kilo\metre} of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}

>>>>

\begin{table}%[tbhp] 
\centering 
\caption{Exposure metrics considered to assess county-level exposure to 
tropical cyclones}
\begin{tabular}{p{1.5cm}p{2.5cm}p{10cm}} 
\toprule
Metric & Available & Criteria for exposure \\ \midrule 
Distance & 1988\,--\,2015 & County population mean center within 100~\si{\kilo\metre} of storm track \\ 
Rain & 1988\,--\,2011 & County cumulative rainfall of $\ge$75~\si{\milli\metre} over the period from two days before to 
one day after the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the 
storm track \\ 
Wind & 1988\,--\,2015 & Modeled storm-associated maximum sustained winds of $\ge$15~\si{\metre\per\second}
at the county's population mean center\\ 
Flood & 1996\,--\,2015 & Flood event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track \\
Tornado & 1996\,--\,2015 & Tornado event listed in the \ac{NOAA} Storm Events database for the county with a start date within two days of the storm's closest approach and county population mean center within 500~\si{\kilo\metre} of the storm track\\
\bottomrule 
\end{tabular} 
\label{tab:exposuremetrics} 
\end{table}
